Resilience and Home-Space for WSTS \thanks{This work was partly done while the authors were supported by the Agence Nationale de la Recherche grant BraVAS (ANR-17-CE40-0028).}

### 2023-10-20 VMCAI-revised-version/final.tex
up

@@

\begin{theorem}\label{downward srp}
{\sc State-resilience} is decidable for $\post$-effective, $\pred$-basis-effective, $\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
and $\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

@@

Strong downward compatibility implies furthermore the decidability
of {\sc $k$-state-resilience} and {\sc bounded-state-resilience}.
\alain{trop rapide}
Indeed, with strong downward compatibility, 
$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is,
and becomes possible to represent \mathieu{... and compute ?} and compare
$\pred^{\leq k}(\Safe)$ 
to other downward-closed subsets.

\begin{corollary}\label{downward brp}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\pred$-basis-effective, $\post$-ideal-effective ???? WSTS, strong downward compatibility and upward compatibility when
$\Safe = \mathop{\downarrow} \Safe$.
\end{corollary}

@@

>>>>

\begin{proof}
\mathieu{preuve résumé / concise à écrire}
\iffalse
\alain{preuve pas claire, à réécrire}
% \alain{les reset-VASS sont-ils effective $\post^*$-basis ?}
% \mathieu{non, sinon on aurait décidabilité pour la state résilience}
%This stems from the fact that it is undecidable for Minsky machines with more than one counter,
%which can be simulated by WSTS.
First, let us show that {\sc Resilience} is undecidable for Minsky machine with at least $3$ counters.
%	This is due to the undecidability of $2$-counter Minsky machine termination~\cite{Min61,Min67}.
From a $2$-counter Minsky machine $M$ with final control-state $q_f \in Q_M$, one can construct a $3$-counter Minsky machine $M'$ with final control-state $q_f' \in Q_{M'}$
such that machine $M$ terminates for all inputs iff machine $M'$ can reach $q_f'(0,0,0)$ from any input in $Q_{M'} \times \mathop{\uparrow} (0,0,1)$. 
%			with at least $1$ on its third counter. 
We build $M'$ to simulate $M$ until it reaches $q_f$, then decrease the first two counters until they reach $0$, then, and only then, finally decrease the third counter until it reaches $0$.
Remark that in the construction, from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$, it is not possible to reach an element in $Q_{M'} \times \{ (1,0,0), (0,1,0),(1,1,0)\}$.
Based on this construction, the problem of deciding whether the downward-closed set $Q_{M'} \times \mathop{\downarrow} (1,1,0)$ is reachable in the $3$-counter Minsky machine $M'$
\alain{review 3: - p. 10: why not choose $\mathop{\downarrow}{(0,0,0)}$ instead of $\mathop{\downarrow}{(1,1,0)}$?}
from any input in $Q_{M'} \times \mathop{\uparrow}(0,0,1)$ is equivalent to the termination of machine $M$, that is undecidable ~\cite{Min61}. 
Hence {\sc Resilience} is undecidable for Minsky machines.
\alain{écrire qqchose ici}
Now, executions of Minsky machines can be simulated by reset-VASS~\cite{araki1976PN,dufourd1998reset}. 
% \alain{\cite{araki1976PN} simule-t-il les Minsky machines et comment ?}
%Reset-VASS extend the basic VASS model with special “reset
%transitions” that set to $0$ some coordinates in the vector. 
%			Let us recall their definition here.
%      	it is well known that reset-VASS are WSTS~\cite{dufourd1998reset}.
The simulation operates by building a reset-VASS in which
there exists
 a sequence from 
an initial state %$(s,x_1 \ldots x_n)$
to
a state $ q(0,  \ldots 0)$ if and only if
there exists
a sequence
 in the Minsky machine
 from a corresponding initial state
 to
 the state $q_f(0,0,0)$
  (from Theorem~$5$ in \cite{araki1976PN}).
Hence $Q_{M'} \times \mathop{\downarrow} (1,1,0)$-resilience hold in the Minsky machine $M'$ iff 
resilience hold in the reset-VASS simulating $M'$.
Since reset-VASS can simulate executions of a Minsky machine, {\sc Resilience} is undecidable for reset-VASS and hence for  $\pred$-basis-effective, 
$\omega^2$-WSTS with strong compatibility as well.\qed
%\alain{déjà dit avant en mieux omega2 et strong comp et $\post$-effective mais reset-VASS apportent effective $\pred$-basis et ideally-effective....}
\fi
\end{proof}

@@

\begin{theorem}\label{downward srp}
{\sc State-resilience} is decidable for $\post$-effective, 
%$\pred$-basis-effective, 
$\post$-ideal-effective and $\pred^*$-basis-effective WSTS with downward and upward compatibilities
and $\Safe = \mathop{\downarrow} \Safe$.
\end{theorem}

@@

In the case of strong downward compatibility,
$\pred^{\leq k}(\Safe)$ is downward-closed when $\Safe$ is.
Hence, 
assuming a $\pred^*$-basis-effective variation
consisting in an algorithm
accepting any ideal $I$ and returning a finite basis of
 $\mathop{\downarrow} \pred(I)$, it becomes possible to perform the same procedures as above except with comparisons against $\pred^{\leq k}(\Safe)$ rather than $\pred^*(\Safe)$, and hence {\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable in this case.

\begin{corollary}\label{downward brp}
{\sc $k$-state-resilience} and {\sc bounded-state-resilience} are decidable for $\post$-effective, $\post$-ideal-effective WSTS, strong downward compatibility and upward compatibility when
$\Safe = \mathop{\downarrow} \Safe$
and there exists an algorithm accepting any ideal $I$ and returning a finite basis of
 $\mathop{\downarrow} \pred(I)$.
\end{corollary}

\mathieu{Donner des exemples dans cette classe.}

@@

\section{Reset-VASS and Minsky-machines}\label{minsk}
